---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
  
```{r, message=F}
library(igraph)
#library(tidyverse)
```

# Network Visualizations

## Aspects of visualizations

> - Nodes
> - Edges
> - Location

## Nodes

> - Information can be conveyed by:
>   - Shape, size, color
> - Shape
>   - Typically categorical (e.g., gender, age range)
> - Size
>   - Often a network measure, but can be something about node
> - Color
>    - Often community detection

## Examples

```{r}
#G = graph_from_data_frame(read.csv('../one_k_redittors.csv'))
G = random.graph.game(15, .4)
plot(G)
```

## With node shapes

```{r}
V(G)$shape = sample(c('square', 'circle','sphere'), size = gorder(G), replace = T)
plot(G)
ages = c('18-25', '26-34', '34-50')
```

## Nodes sized by betweenness centrality

```{r}
V(G)$shape = 'circle'
plot(G, vertex.size = betweenness(G) * 2.5)
```

## Nodes colored by community

```{r}

community = membership(fastgreedy.community(G))
colors = c('red', 'orange', 'darkgreen')
V(G)$color = colors[community]
plot(G)
```

## Edges

> - Size
>   - Typically represent weight of relationship
> - Color
>   - Typically represents different types of relationships

## Examples

## Edge width as weight
```{r}

# This code creates a random network, with two edge types and four weights
# Don't worry about understanding this code - just worry about understanding the visualization part that comes next

make_edges = function(from, nodes){
  x = 1:nodes
  x = x[x != from]
  to = sample(x, 1)
}

nodes = 15
edges = 36
from = sample(1:nodes, size = edges, replace = T)
to = unlist(lapply(from, make_edges, nodes = nodes))
df = data.frame(from, to)
df$type = sample(c('friendship', 'advice', 'hatred'), size=nrow(df), replace = T)
df = unique(df)

G = graph_from_data_frame(df)
#E(G)$type = df$type
E(G)$weight = sample(1:4, size = ecount(G), replace = T)
```

```{r}
E(G)$arrow.size=.5

plot(G, edge.width = E(G)$weight)
```

## Edge color as type


```{r}
# First we choose the colors we want to use
colors = c('gray', 'orange', 'green')
# Then we change the "type" of each edge into a number (Uncomment the commented code to print the types_as_numbers variable to see how this works.)
types_as_numbers = as.numeric(as.factor(E(G)$type))
#print(E(G)$type)
#print(types_as_numbers)

# Then, we select a color from the colors vector we created above, according
# to the type
E(G)$color = colors[types_as_numbers]

plot(G)
```

## Position

> - Can represent literal distance
>   - Cities
>   - Seating chart
> - Or social distance
>   - Formal hierarchy
>   - Degree centrality

## Node placement can make a big difference in how a network is perceived

```{r, out.width='32%'}
# Each of these shows a different layout of the same random graph; note how different the conclusions we might draw are.
G = random.graph.game(15, .4)
plot(G, layout = layout_in_circle(G))
plot(G, layout = layout_randomly)
plot(G, layout = layout_as_star)
plot(G)
plot(G, layout = layout_on_grid)
plot(G, layout = layout_with_lgl)
```


# School Data

Now it's your turn. We are going to have a visualization competition, to see who can make the most interesting and insightful visualizations of a new dataset.

The data comes from a Dutch school class in 2003, and is described [at this website](http://www.stats.ox.ac.uk/~snijders/siena/tutorial2010_data.htm).

In order to make things easier, I've changed this messy data of multiple matrices into an igraph object for you. You can see how I did that [here](https://github.com/jdfoote/Communication-and-Social-Networks/blob/master/activities/knecht_school_data.r).

So, what you'll need to do first is load the data.

```{r}
load(url('https://github.com/jdfoote/Communication-and-Social-Networks/raw/master/activities/school_graph.Rdata'))
```

If this doesn't work, then download the file from that URL, and do `load('/path/to/file')`

There are two different graphs that get loaded. `G` is the full graph, and includes two kinds of edges: whether A and B are friends, and whether they went to the same primary school.

```{r}
# This shows all of the edges and attributes about them
E(G)[[]]
```
The `friend_net` igraph object just includes friendship edges. I created it by doing the following:

```{r}
# This makes a "subgraph" by just getting the edges with the type 'friendship'
friend_net = subgraph.edges(G, E(G)[E(G)$type=='friendship'])
```

Both of these graphs include lots of data about the people in them. You can view the attributes of a vertex like this:

```{r}
V(friend_net)[[]]
```
Descriptions of what each measure means are at the [site the data came from](http://www.stats.ox.ac.uk/~snijders/siena/tutorial2010_data.htm).

Basically, it includes information about alcohol use, gender, age, ethnicity (whether Dutch or not), and religion.

## The Goal

Your assignment during class is to come up with something interesting that a network visualization of this data could tell us, and a drawing (on paper) of what you want it to look like.

## Homework

Your homework (due the Tuesday after Spring Break) is to create a beautiful, insightful visualization of this network that helps us to see something that might not otherwise be apparent.
