---
title: "Ego Network Calculations"
author: "Your Name Here"
format: pdf
---

```{r setup, message=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# import the libraries needed
library(igraph)
library(ggraph) # If you get an error run install.packages('ggraph')
library(tidygraph) # If you get an error run install.packages('tidygraph')
library(readxl) # If you get an error run install.packages('tidyverse')
#extrafont::loadfonts() 
set_graph_style(family="sans")
```

Create your networks. Make an node attributes file for each person that you have "discussed important matters with" over the last 6 months.

The file should include:
- sex
- education (years of education)
- race
- age
- is.kin (are they related to you?)

Then, for each person, make an edge list of which of the other people they are connected to. For each edge, put the edge weight as 0 if they are strangers, 1 if they are especially close, or .5 if they are somewhere in between.

```{r}
edges <- read_excel('~/Teaching/communication_and_networks/activities/ego_networks/ego_edges.xlsx')
nodes <- read_excel('~/Teaching/communication_and_networks/activities/ego_networks/ego_atts.xlsx')
```

Code to load the file into R and make it into a network object.

```{r}
G <- graph_from_data_frame(d=edges, vertices = nodes, directed = F) %>% as_tbl_graph()
```

Code to filter out the edges that have a weight of 0.

```{r}
G <- G %>%
  activate(edges) %>%
  filter(!is.na(weight)) %>%
  filter(weight != 0)
```

This code will make a plot of the network. You'll learn soon how to do this on your own.

```{r}
G |>
  ggraph(layout = 'stress') +
  geom_edge_link(aes(width = weight), alpha = .5) +
  geom_node_point(aes(color = is.kin), size = 5) +
  geom_node_text(aes(label = name), repel = TRUE, size = 8) +
  scale_color_viridis()
```


## Exercises

Calculate the following for your network. Use Google to find the code to do this. I'll give you the first one.

1. The number of nodes

```{r}
G |>
  vcount()
```

2. The number of nodes who are kin (use the `filter` function)
```{r}

## YOUR CODE HERE


```

3. The number of nodes who are non-kin

```{r}

## YOUR CODE HERE

```


4. The density of the network
```{r}

## YOUR CODE HERE

```