---
title: COM 411
template: "../resources/template.html"
output: 
  revealjs::revealjs_presentation:
    theme: white
    progress: true
    transition: convex
    self_contained: false
    reveal_plugins: ["notes"]
    lib_dir: ../../libs
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = './')
source("resources/preamble.R")

f <- function (x) {formatC(x, format="d", big.mark=',')}
bold <- function(x) {paste('{\\textbf{',x,'}}', sep ='')}
gray <- function(x) {paste('{\\textcolor{gray}{',x,'}}', sep ='')}
wrapify <- function (x) {paste("{", x, "}", sep="")}
p <- function (x) {formatC(x, format='f', digits=1, big.mark=',')}
library(igraph)
```

<style type="text/css">

/* Changes to RevealJS default formatting.

I prefer this style but you can change these to your preferences */

/* This bit lets you do columns */
.container{
  display: flex;
}

.col{
  flex: 1;
}

.reveal section img{
border: none;
background: none;
box-shadow: none;
}

.highlight {
color: LightCoral;
}

.reveal h1,
.reveal h2,
.reveal h3,
.reveal h4,
.reveal h5,
.reveal h6 {
text-transform: none;
}

.reveal h1 {
margin-top:200px;
}

div.footnote {
font-size: 40%;
text-align: right;
}

.bg-image{
width: 180px;
}

.reveal .slides section .fragment.fade-in-then-out,
.reveal .slides section .fragment.current-visible {
  opacity: 0;
  visibility: hidden; }
  .reveal .slides section .fragment.fade-in-then-out.current-fragment,
  .reveal .slides section .fragment.current-visible.current-fragment {
    opacity: 1;
    visibility: inherit; }

.reveal .slides section .fragment.fade-in-then-semi-out {
  opacity: 0;
  visibility: hidden; }
  .reveal .slides section .fragment.fade-in-then-semi-out.visible {
    opacity: 0.5;
    visibility: inherit; }
  .reveal .slides section .fragment.fade-in-then-semi-out.current-fragment {
    opacity: 1;
visibility: inherit; }

</style>

## Today's Dad Joke

Shout out to people wondering what the opposite of 'in' is.


## Housekeeping

> - Reading academic articles
>	- Don't read everything
> - Etherpad
>	- How to run notetaking?
>	- Collaborative summaries during the week?

## Schedule

> - Housekeeping / Announcements (5-10)
> - Discussion and review (40-45)
> - Consolidation and Confusion (10-15)
> - Supplementary lectures
>	- Discuss at end of class





# Network Data and Network Types



## Homework

>- Did you find someone with a Hawaiian driver's license?
>- How did it go?
>- What were the challenges?
>- Were there overlapping networks?

# Networks in R

## What is R?

## Why are we using R?

>- Powerful
>- Reproducible
>- Extensible


## Network Data

>- What are the three main ways of representing networks?

## Matrices

```{r, include = T}
sw = watts.strogatz.game(2, 4, 1, .2)
as_adjacency_matrix(sw)
```

## Edgelists

```{r, include = T}
as_edgelist(sw)

```

## Graphs/plots/sociograms

```{r, include = T}
plot(sw,
     vertex.label.cex = .5
     )
```

## Where does social network data come from?

>- Surveys
>- Observation
>- Trace data

# Network Types

## Ego networks

>- Typically created from surveys
>- Your family networks were ego networks

```{r, include = T}
plot(make_ego_graph(sw, nodes = 1)[[1]])

```

## Extended ego networks

```{r}
plot(
  make_ego_graph(sw, 
                 nodes = 1, 
                 order = 2)[[1]])

```

## Bipartite Networks

>- 20 random reddittors and their communities
```{r, echo = F, message=F, echo=F}
library(tidyverse)
r = read_csv('../../resources/one_k_redittors.csv')
r = r %>% filter(author %in% sample(unique(r$author), 50))
r = r %>% filter(! subreddit %in% r$author)
G = graph.data.frame(r, directed = T)
V(G)$type = TRUE
V(G)[name %in% r$subreddit]$type = FALSE

l = layout.bipartite(G, )
colors = c('steelblue', 'orange')
shapes = c('circle', 'square')
```
```{r, echo=F}

plot_graph = function(g){
  plot(g,
     layout = layout.fruchterman.reingold(g),
     #vertex.color = colors[V(g)$type + 1],
     #vertex.shape = shapes[V(g)$type + 1],
     vertex.size = 4,
     vertex.label.cex = .2
       )
}

G %>% induced_subgraph(unlist(sapply(1:20, subcomponent, graph = G, mode = 'out'))) %>%
  plot(
     layout = layout.bipartite(.),
     vertex.color = colors[V(.)$type + 1],
     vertex.shape = shapes[V(.)$type + 1],
     vertex.size = 4,
     vertex.label = ""
     #edge.width = log1p(E(G)$posts)
     )
```

## What does this data look like?

```{r}
head(r)
```

## Bipartite networks can be "projected"

>- People who comment in the same communities


```{r}
projections = bipartite_projection(G)
sr = projections[[1]]
people = projections[[2]]

people %>% delete.vertices(degree(.) < 1) %>%
  plot_graph()

```

## Projections

```{r}
sr %>% delete.vertices(degree(.) < 1) %>%
  plot_graph()
```

## Nodes and edges can have attributes

```{r}
V(sw)$gender = sample(c(1,2,3), 
                      length(V(sw)), 
                      replace = T)

colors <- c('green','orange','red')
V(sw)$color <- colors[as.integer(V(sw)$gender)]
plot(sw
     )
```

## Edge attributes

>- Usually this is weight

```{r}
#tmp = delete.edges(sr, which(E(sr)$weight < 2))
people %>%
  delete.vertices(degree(.) < 1) %>%
  plot(
    layout = layout_with_fr(.),
    vertex.size = 9,
    vertex.label.cex = .4,
    edge.width = E(.)$weight
  )
```


## Assignments

> - Homework 3 (extended family network)
> - Install R and RStudio (optional)


# Intro to R

## The Dad Joke

Shout out to people wondering what the opposite of 'in' is.

## Homework

> - How did it go?
> - What was different from the first assignment?

## Questions about R?

> - We will regroup at the end of class
> - We can adjust how quickly we go through things

## For Tuesday

Turn in Self-Assessment Reflection (instructions on wiki)
Finish Introduction to R?

