---
title: "Business Example"
author: "Jeremy Foote"
date: "4/15/2020"
output: html_document
---

For the final project, you might want to use some data from an actual company to showcase the sort of analysis that you could do for the organization that you choose to focus on.

There is a great set of network data at https://github.com/schochastics/networkdata. You can install all of them with the following (you will only need to run this once - uncomment these lines, run them, and after that comment them out with # like I do below).

```{r}
#install.packages('drat')
#drat::addRepo("schochastics")
#install.packages("networkdata")
```
After it's installed, we load it and set options

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(networkdata)
library(igraph)
```

The data we want is stored in three igraph networks that came from observing and surveying managers in a high-tec company. They are called `ht_advice`, `ht_friends`, and `ht_reports`.

There are a few other data sources that might be interesting to you - you can look at all of them by running `data(package = "networkdata")`


Here are a few functions you might want to use as you prepare your final projet. This is one way to prepare a report about who is influential.

```{r}
# Get and store centrality measures in the vector "spreadsheet"
V(ht_advice)$bw = betweenness(ht_advice, directed = T)
V(ht_advice)$degree = degree(ht_advice, mode = "in")
V(ht_advice)$closeness = closeness(ht_advice, mode = "in", normalized = T)

# Then, print out just the columns we want
df = as.data.frame(vertex_attr(ht_advice))
df$node_number = 1:21
df[,c(8,5:7)]
```

Or maybe you want to resize the nodes by one of these measures of power/influence.

```{r}

E(ht_advice)$arrow.size = .25

V(ht_advice)$size = V(ht_advice)$closeness * 30 # You may need to change what you multiply it by for different measures
  
plot(ht_advice)

```

Or you might look at whether younger members are being integrated... (remember, you can view attributes using this code):

```{r}
V(ht_advice)[[]]
```


For example, this shows those under the median age as a different color, and uses the friendship network.

```{r}
V(ht_friends)$color = 'tomato'
V(ht_friends)[which(age < median(age))]$color = 'royalblue'
plot(ht_friends, edge.arrow.size = .25)
```

Finally, we might look for cross-departmental ties. Here's one way to do that:

```{r}
# Set a color for each department (1-4)
dept_colors = c('orange','green','tomato','royalblue')

# Set the color of each node based on the department
V(ht_advice)$color = dept_colors[V(ht_advice)$dept]

plot(ht_advice)

```

