---
title: Week 13 - Collective behavior
template: "../resources/template.html"
output: 
  revealjs::revealjs_presentation:
    theme: white
    progress: true
    transition: convex
    self_contained: false
    css: ../../libs/custom_css.css
    reveal_plugins: ["notes"]
    lib_dir: ../../libs
    allow_uptree_lib_dir: True
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = './')
source("resources/preamble.R")
set.seed(0121)
library(igraph)
library(tidygraph)
library(ggraph)
library(tidyverse)
df = read_csv('~/Teaching/communication_and_networks/week_13/lecture/wisdom_of_crowds_activity.csv')
```

## Dad Joke

Never date a tennis player.

<p class='fragment'>
Love means nothing to them.
</p>

## Housekeeping

> - Exam is due on Thursday before class
> - Under quizzes on Brightspace
> - Remember, it will give you new random questions if you come back
> - For the R question, turn in what you get even if you don't get the whole thing right
> - Wisdom or Madness of Crowds on Thursday
> - Apologies about schedule confusion!!!
> - No class on Tuesday - discussion on Discord
>   - Try to be online during class time


## Upcoming Assignments

> - R Lab 8 - November 28
> - Final project rough draft (December 1)
> - Peer feedback (Due December 6)

## Review Questions

> - What is "collective behavior"?
> - What does collective behavior have to do with networks?
> - What is "the wisdom of crowds"?
> - What is groupthink? Have you seen examples?
> - In Dr. Becker's experiments, why did groups get better when communicating?

## Wisdom of Crowds Activity

> - Get out a piece of paper
> - Guess how many pages are in this book and write down your guess
> - Tell your guess (and no other information) to the two people on either side of you
> - Guess again how many pages are in the book and write it on a new line (Guess 2)

## Activity 2

> - Guess 3: Guess the pages in this book
> - Two random people will tell us their guess
> - Guess 4: Guess the pages again for this book


## Results

## Decentralized network

```{r}
df %>%
  filter(Network_Type == 'Decentralized') %>%
  ggplot() +
  geom_density(aes(x = Guess, fill = as_factor(Round)), alpha = .3) + #, position='dodge', bins = 5) + 
  geom_vline(xintercept=294, color = 'red', linetype='dashed') + 
  theme_minimal() + 
  scale_fill_viridis_d(name='Round')
```

## Centralized network

```{r}
df %>%
  filter(Network_Type == 'Centralized') %>%
  ggplot() +
  geom_density(aes(x = Guess, fill = as_factor(Round)), alpha = .3) + #, position='dodge', bins = 5) + 
  geom_vline(xintercept=377, color = 'red', linetype='dashed') + 
  theme_minimal() + 
  scale_fill_viridis_d(name='Round')
```

## Wisdom and Madness of Crowds

## Questions

> - What is the "majority illusion"?
> - How does the majority illusion relate to the friendship paradox?
> - Majority illusion from Nicky Case activity?

## Discussion Questions

