---
title: Week 9
template: "../resources/template.html"
output: 
  revealjs::revealjs_presentation:
    theme: white
    progress: true
    transition: convex
    self_contained: false
    css: ../../libs/custom_css.css
    reveal_plugins: ["notes"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = './')
source("resources/preamble.R")
set.seed(0121)
library(igraph)
library(tidygraph)
library(ggraph)
library(tidyverse)
```

## Dad Joke

Why did the raisin go out with the prune?

<span class='fragment'>
Because he couldn't find a date. 
</span>

## Housekeeping

> - Video for next week coming
> - Thursday - 6 Degrees of Wikipedia + R Review
> - Self-assessment reflections
> - Visualization challenge is due next week
>	- "Knit" to Word Doc
> - Light reading

## Review Questions

> - What is a degree distribution?

<div style="position:relative; width:640px; height:480px; margin:0 auto;">
<div class='fragment' style="position:absolute;top:0;left:0;">
```{r}
set_graph_style()

G = play_barabasi_albert(10000, growth = 1, power = 1)

df = G %>% 
  as_tbl_graph() %>%
  activate(nodes) %>%
  mutate(degree = centrality_degree(mode='all')) %>%
  as_tibble() %>%
  group_by(degree) %>%
  tally(name = 'count')


df %>% 
  filter(degree > 1) %>%
  ggplot(aes(x = degree, y = count)) + 
  geom_line() + 
  theme_minimal() +
  labs(title="All nodes")

```

</div>

<div class='fragment' style="position:absolute;top:0;left:0;">
```{r}
df %>% 
  filter(degree > 2) %>%
  ggplot(aes(x = degree, y = count)) + 
  geom_line() + 
  theme_minimal() + 
  labs(title="Nodes with degree of at least 3")
```
</div>

<div class='fragment' style="position:absolute;top:0;left:0;">
```{r}
df %>% 
  filter(degree > 3) %>%
  ggplot(aes(x = degree, y = count)) + 
  geom_line() + 
  theme_minimal() + 
  labs(title="Nodes with degree of at least 4")
```

</div>
</div>

## Review Questions

Think, Pair, Share

> - What is a scale-free distribution?
> -	When/where do we find scale-free distributions?
> - What is an example of a network that would *not* have a scale-free distribution?
> - What are the benefits of scale-free networks?

## Review Questions

Think, Pair, Share

> - What is the "friendship paradox"?
> - How is the friendship paradox useful?


## Discussion Questions


## Visualization Challenge

> - Similarity visualization


## Work on Tidygraph Tutorial
