---
title: Week 11 - Communities in Networks
sort: 5
format: 
  revealjs:
    css: /resources/custom_css.css
    theme: white
    progress: true
    transition: convex
    reveal_plugins: ['notes']
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = './')
set.seed(0121)
library(igraph)
library(tidygraph)
library(ggraph)
theme_set(theme_graph()) 

library(igraphdata)
data("UKfaculty")

UK <- as_tbl_graph(UKfaculty)

UK <- UK |>
  activate(edges) |>
  filter(edge_is_mutual()) |>
  to_undirected() |>
  simplify() |>
  as_tbl_graph()
```

## Dad Joke

The guy who stole my diary died yesterday.

<span class='fragment'>
My thoughts are with his family.
</span>


## Housekeeping

> - Make an appointment for Self Assessment Reflection review
> - Submit Exam Questions ASAP

## Plan for today

> - Groups in networks review
> - Discussion questions
> - R

<!---
# Troubled Lands Review

## Questions from Brightspace Assignment

> - How did you play differently in different versions?
> - What in real life is similar to each version of the game?
> - How did you decide whether or not to help others?
> - What does this have to do with social networks and social capital?!


## Debrief

> - What did you think about while playing?
> - How did you feel as you played? Did those feeling change across different versions of the game?
> - When do you help others in real life?

## Debrief

> - Were resources and abilities distributed fairly?
> - How should resources be distributed in the real world?
> - What role did communication play? How did your group make decisions?
> - Did you work to make things equal? Why or why not?
> - What kinds of inequality appeared in the game?
--->

## Community Detection Review

> - What is a component? (Hint: there's often one "giant" component)
> - What is a clique?
> - What is the weakness of cliques as a way of identifying groups?
> - What is a k-plex? (Hint: related to cliques)
> - The goal of most community detection algorithms is high modularity. What does that mean?

## Core and Periphery Review

> - Why do groups often have a core and a periphery?
> - Given some of the other theories we've talked about, how would you expect the periphery to differ from the core?
> - What role does the periphery have in protests (Barberá et al. reading)?
> - What is one way to identify who is in the core or periphery?

## Coreness Example

<div class='col fragment'>
```{r,message=F}
UK |>
  activate(nodes) |>
  mutate(coreness = node_coreness()) |>
ggraph(layout='stress') +
  geom_edge_fan(alpha=.3) +
  geom_node_point(aes(color=coreness)) +
  scale_color_viridis_c(name='Coreness')

```
</div>

## Discussion Questions
