---
title: Week 12 - Collective Behavior
format: 
  revealjs:
    css: /resources/custom_css.css
    theme: white
    progress: true
    transition: convex
    reveal_plugins: ['notes']
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = './')
source("resources/preamble.R")
set.seed(0121)
library(igraph)
library(tidygraph)
library(ggraph)
library(tidyverse)
library(networkdata)
# Set graph theme
df = read_csv('./resources/wisdom_of_crowds_activity.csv')

```

## Today's Dad Joke

When I was young, we were really poor. My family could only afford a second-hand calculator, which was missing the "X" button.

<p class='fragment'>
Times were hard.
</p>

## Housekeeping

> - Exam next week
>	- Open book / open note / open internet?
>	- Short response  + multiple choice
> - Review on Thursday

## Final Project

> - Timing for final project
>	- Don't wait - get started now!
>	- Reach out if you get stuck
> - Look at example projects from the past
> - Not all visualizations have to be networks
> - Concepts + Visualizations

# Mini-lecture on Wisdom of Crowds and Collective Intelligence

## Collective Intelligence

> - Idea that groups can be smarter than the smartest person in the group
> - Sometimes it's simple; e.g, aggregating information from multiple sources
> - Sometimes there is an emergent property, where the group thinks and behaves differently

## Ants

{{< video https://www.youtube.com/watch?v=IQW_K8rq1N4 height="85%" width="100%" >}}

## Organizations

> - While organizations are much more hierarchical than ants, they can still have emergent properties
> - Partially a function of history, partially of the structure of the organization
> - Even a culture everyone dislikes can be hard to change
> - E.g., metric system, QWERTY keyboards

## Cities

> - High-level decisions about zoning, infrastructure, etc. but the character of a city is emergent
> - Jane Jacobs: "The Death and Life of Great American Cities"


## Wisdom of Crowds Activity

> - Get out a piece of paper
> - Guess how many pages are in this book and write down your guess
> - Tell your guess (and no other information) to the two people on either side of you
> - Guess again how many pages are in the book and write it on a new line (Guess 2)

## Activity 2

> - Guess 3: Guess the pages in this book
> - Two random people will tell us their guess
> - Guess 4: Guess the pages again for this book


## Collective understanding and decision-making

> - Wisdom of crowds is idea that in many contexts, the average is a really good guess!
> - What is groupthink? Have you seen examples? How might this relate?
> - Dr. Josh Becker et al. show that in some cases, communication can help
>   - Those whose guesses are worst move the most
>   - Divorced from other social relationships (not multiplex)


## Results

## Decentralized network

```{r}
df |>
  filter(Network_Type == 'Decentralized') |>
  ggplot() +
  geom_density(aes(x = Guess, fill = as_factor(Round)), alpha = .3) + #, position='dodge', bins = 5) + 
  geom_vline(xintercept=294, color = 'red', linetype='dashed') + 
  theme_minimal() + 
  scale_fill_viridis_d(name='Round')
```

## Centralized network

```{r}
df |>
  filter(Network_Type == 'Centralized') |>
  ggplot() +
  geom_density(aes(x = Guess, fill = as_factor(Round)), alpha = .3) + #, position='dodge', bins = 5) + 
  geom_vline(xintercept=374, color = 'red', linetype='dashed') + 
  theme_minimal() + 
  scale_fill_viridis_d(name='Round')
```

## Connections to other network concepts

> - How does this relate to our week on small group networks?
>   - Transactive memory
>   - Hierarchy
>   - Centralization
> - How could you design a group / network to get the benefits of the wisdom of crowds?
> - When would you expect a wisdom of crowds approach to work well?
> - When would you expect it to fail?

## Plan for Today

> - Exam Review


## Exam Review

>	- Visualization / ggraph
>   - Look at code and explain what it will look like
>   - Find the bug in code

## Example

> - What code produces this graph?

<div class='fragment'>

```{r, echo=FALSE}
G <- play_gnp(18, .15, directed = F) |> as_tbl_graph()
set_graph_style(family='sans')

G |>
  mutate(centrality = centrality_degree()) |>
  activate(edges) |>
  ggraph() +
  geom_edge_fan(color = 'purple') +
  geom_node_point(aes(size = centrality), color = 'green')
```

</div>

<div class='fragment'>

```r
G |>
  mutate(centrality = centrality_degree()) |>
  activate(edges) |>
  ggraph() +
  geom_edge_fan(color = 'purple') +
  geom_node_point(aes(size = centrality), color = 'green')
  ```

</div>

## Exam Review

What would this code do?

```r
G |>
  activate(nodes) |>
  mutate(ec = centrality_eigen()) |>
  filter(ec > 1) |>
  ggraph() +
  geom_edge_fan(color = 'orange') +
  geom_node_point(aes(size = ec), color = 'green')

```

## Exam Review

> - Centrality
> - Network representations (Edgelists and matrices)
> - Network formation
>   - Homophily, focus theory, transitivity, social exchange theory, etc.
> - Tie strength and small worlds

## Exam Review

> - Power and social capital
> - Contagion and diffusion
> - Centrality
> - Resilience
> - Bipartite networks and projections

## Exam Review

> - Friendship paradox
> - Strength of weak ties

<!--
## Review Questions

> - What is the difference between a filter bubble and an echo chamber?
> - How do echo chambers relate to social influence?
> - What did Chris Bail et al.'s research find about breaking out of echo chambers?
> - How has the internet changed social networks? What are the most important positives and negatives?


## Review Questions

> - What is context collapse?
> - How does context collapse relate to network principles we have discussed?

## Discussion Questions
-->