---
title: Network Visualization and Wrap Up
format: 
  revealjs:
    css: /resources/custom_css.css
    theme: white
    progress: true
    transition: convex
    reveal_plugins: ['notes']
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_knit$set(root.dir = './')
source("resources/preamble.R")

f <- function (x) {formatC(x, format="d", big.mark=',')}
bold <- function(x) {paste('{\\textbf{',x,'}}', sep ='')}
gray <- function(x) {paste('{\\textcolor{gray}{',x,'}}', sep ='')}
wrapify <- function (x) {paste("{", x, "}", sep="")}
p <- function (x) {formatC(x, format='f', digits=1, big.mark=',')}
library(igraph)
library(tidyverse)
library(tidygraph)
library(ggraph)
theme_set(theme_graph())
```

## Today's Dad Joke

Within minutes, the detective knew exactly what the murder weapon was.

<span class='fragment'>
It was a brief case.
</span>

## Housekeeping

> - Peer review
> - Final project due Dec. 13
> - Self Assessment Reflection due Dec. 13
> - Advanced visualizations
>   - Faceting
>   - Grouping
>   - Visualizing homophily
>   - Changing color palettes
>   - Changing/adding titles and legends

# Network Visualizations

## What are network visualizations good for?

> - Uncover something about the structure of the network
>   - And usually how it relates to the attributes of its members

## Aspects of visualizations

> - Nodes
> - Edges
> - Location

## Nodes

> - Information can be conveyed by:
>   - Shape, size, color
> - Shape
>   - Typically categorical (e.g., gender, age range)
> - Size
>   - Often a network measure, but can be something about node
> - Color
>    - Community detection or node attribute or network measure

## Examples

```{r}
#G = graph_from_data_frame(read.csv('../one_k_redittors.csv'))
G = random.graph.game(15, .4) |> as_tbl_graph()
G = G |> activate(nodes) |>
mutate(age = sample(c(19,20,21,22,23), 15, replace=T)) |>
mutate(major = sample(c('Communication', 'Computer Science', 'Sociology'), 15, replace=T))

G |>
ggraph(layout='stress') + 
geom_edge_fan(width=.5) + 
geom_node_point(size = 5)
```

## With node shapes

```{r}
G |>
ggraph(layout='stress') + 
geom_edge_fan(width=.5) + 
geom_node_point(aes(shape=major), size = 5)
```

## Nodes sized by betweenness centrality

```{r}

G |>
mutate(bc = centrality_betweenness()) |>
ggraph(layout='stress') + 
geom_edge_fan(width=.5) + 
geom_node_point(aes(size = bc))

```

## Nodes colored by community

```{r}
G |>
mutate(community = as_factor(group_fast_greedy())) |>
ggraph(layout='stress') + 
geom_edge_fan(width=.5) + 
geom_node_point(aes(color = community), size=5) +
scale_color_viridis_d()
```

## Edges

> - Size
>   - Typically represent weight of relationship
> - Color
>   - Typically represents different types of relationships

## Examples

## Edge width as weight
```{r}

G = G |>
activate(edges) |>
mutate(weight = sample(1:20, nrow(.E()), replace=T),
      type = sample(c('friendship', 'hatred'), nrow(.E()), replace=T))

G |>
ggraph(layout='stress') + 
geom_edge_fan(aes(width=weight)) + 
geom_node_point(size=5)

```

## Edge color as type


```{r}
G |>
ggraph(layout='stress') + 
geom_edge_fan(aes(width=weight, color=type)) + 
geom_node_point(size=5)

```

## Position

> - Can represent literal distance
>   - Cities
>   - Seating chart
> - Or social distance
>   - Formal hierarchy
>   - Degree centrality

## Node placement can make a big difference in how a network is perceived

```{r, out.width='32%'}
G = play_islands(3, 5, .7, 2)
G |>
ggraph(layout='stress') + 
geom_edge_fan(width=.5) + 
geom_node_point(size=5)

G |>
ggraph(layout='circle') + 
geom_edge_fan(width=.5) + 
geom_node_point(size=5)

G |>
ggraph(layout='randomly') + 
geom_edge_fan(width=.5) + 
geom_node_point(size=5)

G |>
ggraph(layout='fr') + 
geom_edge_fan(width=.5) + 
geom_node_point(size=5)
```

## What makes for a good network visualization?

Kozo Sugiyama's rules:

<div style='font-size:smaller'>

1. Lines should be straight.
2. Lines should be far apart from one another.
3. Lines should not cross or touch.
4. Lines should be easy to follow from one node to another.
5. Nodes that connect should be close.
6. Nodes that are most central should be close to the center of the graph.
7. Nodes that are similar in some way should be placed near one another.

</div>

## Activity

Find and Assess a network visualization

> - Use Google Images (or similar) to find an image of a network graphs.
> - What do you think the image does well and poorly? Does it follow Sugiyama's rules?
> - Put the image (and your assessment) on the padlet at https://padlet.com/jdfoote1/networks (link is also on wiki)

## Discussion

> - Did you agree with Sugiyama's rules?
> - What types of color/shape/size approaches were effective?
> - Did this make you want to change any visualizations for your final project?

## Reflection

> - What concepts have you found most compelling?
> - How has the class changed your perspective?

## Parting thoughts

> - "We shape our [networks]; thereafter they shape us" - Churchill, paraphrased
> - We are embedded in mostly invisible networks of influence
>   - Our beliefs, opportunities, and outcomes are not independently chosen or earned
>   - The choices we make may extend far beyond our vision
> - Cultivate our social capital for good
> - I am grateful for the edges created during our class. :)
