---
title: "Groups in Networks"
author: "Jeremy Foote"  
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(igraph)
library(ggraph)
library(tidygraph)
```

In this lesson, we will focus on understanding different approaches to identifying groups in networks. Human networks are "groupy" - whether through triadic closure or foci of activity, and so they typically have groups. If you think about your own friendship network, it's likely that you could group many of them (e.g., high school friends, soccer friends, and Comm friends). Network structure can be used to identify different groups.

## Data

For much of this lesson, we will use the `UKfaculty` frienship data which is part of the `igraphdata` package.

Let's load it as `UK` and take a quick look at it

```{r}

#install.packages('igraphdata') # Run this if it isn't installed yet
library(igraphdata)
data("UKfaculty")

UK <- as_tbl_graph(UKfaculty)

UK %>%
  activate(edges) %>%
  ggraph() +
  geom_edge_fan(alpha=.4) +
  geom_node_point(aes(color=as.factor(Group))) +
  scale_color_viridis_d(name="Department")
  
```


Whenver loading a new dataset, it's worth looking at it. Because we loaded this as a tidygraph `tbl_graph`, we can just type the name of the graph, and it will show the node and edge attributes.


```{r}
UK
```
So, for nodes we know the department that each person belongs to (the "Group" node attribute) and a weight representing the strength of the relationship. Some of the algorithms we are going to be working with assume undirected networks, so we're going to filter this network to only mutual edges, and treat it as undirected.

```{r}
UK <- UK %>%
  activate(edges) %>%
  filter(edge_is_mutual()) %>%
  to_undirected() %>%
  simplify() %>%
  as_tbl_graph()

UK %>%
  ggraph() +
  geom_edge_fan(alpha=.4) +
  geom_node_point(aes(color=as.factor(Group))) +
  scale_color_viridis_d(name="Department")
```


We can use the known group to compare different approaches for identifying groups in networks.

# Approaches to identifying groups in networks

## Components

The simplest version of a group in a social network is all of the people who are connected to each other. This is called a component. Most real-world social or communciation networks include almost everyone in a single, "giant component". This seems strange at first, but an example can help to show why it works this way. Let's say you had a network of 50 people with two large components of 25 people each. If any one of the 25 people in one component adds an edge to any one of the 25 people in the other component, then the components merge. In other words, there are so many possible edges that would combine the components that one of them is bound to occur unless there is something separating the groups.

The following visualizations show a random graph as it goes from 3% density to 15% density. It quickly becomes a single, large component.

This attribute of networks means that components are rarely very interesting as a way of grouping nodes.

```{r}
# Don't worry about understanding this code - it's basically just plotting 
# different densities of random graphs
for(p in c(.02, .05, .08, .10)){
  G = random.graph.game(50, p) %>% as_tbl_graph()
  plot <- G %>%
    ggraph(layout='stress') +
    geom_edge_fan(alpha = .5) +
    geom_node_point() +
    labs(title=paste0("Random graph with ", p*100, "% of edges"))
  print(plot)
}
```


## Cliques

The next simplest idea is a clique - this identifies groups of nodes where everyone is connected to everyone else in the group.

Let's find the largest cliques in this network, using the `largest_cliques` function

```{r, warning=F}

# Find the largest clique(s)
UK %>% 
  largest_cliques()
```
This network has a bunch of cliques of size 6 - that means that there are edges between each pair of people for all 9 of the people in the clique.

Let's visualize this by coloring the first of these cliques.

```{r, warning=F}
# First, get the list of nodes in the first clique
clique_nodes <- largest_cliques(UK)[[1]]
UK %>% 
  activate(nodes) %>%
  mutate(idx = 1:nrow(.N())) %>% # Create an index for the nodes
  mutate(in_clique = idx %in% clique_nodes) %>% # If the index is one of the clique nodes
ggraph(layout = 'stress') +
  geom_edge_fan(alpha=.3) +
  geom_node_point(aes(color=in_clique)) + # Color by whether the node is in the clique
  scale_color_viridis_d(name='In largest clique')
```


This is not a super intuitive way of identifying groups and it misses some clear areas that we would probably consider groups. For example, in the figure above, only a subset of what is clearly a group are actually included in the clique.

A number of tweaks to this approach exist, such as n-cliques (where everyone in the group has to be connected within a distance of n) and k-plexes (where everyone in the group has to be connected to all but k other members). These do help make groups more natural. Unfortunately, neither of these have implementations in igraph so I won't visualize them.

## K-cores

Another approach is called k-cores. This isn't quite the same as identifying the group a node is in; the idea is to identify "central" and "peripheral" parts of a network. Each node gets a "coreness" number, which is the largest k for which it exists in a subgraph where all nodes are connected with at least k edges. This sounds super confusing, but it makes a bit more sense when plotted. Here is some sample code for visualizing coreness.

```{r}

UK %>%
  activate(nodes) %>%
  mutate(coreness = node_coreness()) %>%
ggraph(layout='stress') +
  geom_edge_fan(alpha=.3) +
  geom_node_point(aes(color=coreness)) +
  scale_color_viridis_c(name='Coreness')

```

## Community detection

There are a number of more complicated algorithms for finding communities in networks, often called "community detection" or "clustering". The very bottom of [this page](https://tidygraph.data-imaginist.com/reference/index.html) shows those which have been implemented in ggraph.

Many of them are based on one of two ideas - the first is [modularity](https://en.wikipedia.org/wiki/Modularity_(networks)). These approaches attempt to find partitions of a network which maximize the number of edges within groups and minimize the number of edges between them.

The second approach is based on "random walks". The idea is that if you start at a random place in a network and move along a random edge, and keep doing this over and over, then walks should generally stay within groups and rarely go between them.

I'll show you how to visualize a couple of these in ggraph

First, a modularity approach, using the "fast greedy" algorithm
```{r}
UK %>%
  activate(nodes) %>%
  mutate(community = as.factor(group_fast_greedy())) %>%
ggraph(layout = 'stress') +
  geom_edge_fan(alpha=.3) +
  geom_node_point(aes(color=community)) +
  scale_color_viridis_d(name='Fast Greedy Communities')
```

As you can see, this is pretty good. It groups things pretty similarly to the actual departments.

Here is a label propagation approach, which is similar to random walks:

```{r}
UK %>% as_tbl_graph() %>%
  activate(nodes) %>%
  mutate(community = as.factor(group_louvain())) %>%
ggraph(layout = 'stress') +
  geom_edge_fan(alpha=.3) +
  geom_node_point(aes(color=community)) +
  scale_color_viridis_d(name='Louvain Communities')
```

In most networks, community detection approaches give generally similar results.


# Assignment

1. Change the layout of this graph to a few of the others found at the bottom of [this page](https://igraph.org/r/doc/layout_with_kk.html). See if you can find one that you think does a better job of showing group structure.

2. I have used coloring of nodes as a way to identify different groups in a network. I think that this is a good way of visualizing groups, but there are other options. Come up with a different way of visualizing groups and either draw a picture of your idea or actually code it in ggraph.

3. Go back to your Dutch school network visualization and add a visualization of group structure.
