---
title: Power, centrality, and hierarchy
format: 
  revealjs:
    css: /resources/custom_css.css
    theme: white
    progress: true
    transition: convex
    reveal_plugins: ['notes']
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message=F)
knitr::opts_knit$set(root.dir = './')
source("resources/preamble.R")
library(igraph)
library(tidygraph)
library(ggraph)

set_graph_style('theme_graph')
```


## Today's Dad Joke
Bro, can you pass me that leaflet?


Brochure.

# Power in Social Networks

## What is power?
> - Long, varied history in the social sciences
>   - One version is the capacity to achieve one's will
>   - How do we do that? Through capital
> - Social networks are one way of measuring "social capital"
> - Who has power in a social network?

## Degree Centrality

>- Counts the number of edges each node has
>  - In, out, or all


```{r, out.width="45%", message=F}
set.seed(23)
G = random.graph.game(8, .35)
G <- G |> as_tbl_graph() |>
  mutate(name = LETTERS[1:8])

G <- create_notable("Zachary") |>
  mutate(name = 1:34)

 G |>
  ggraph() +
  geom_edge_fan() +
  geom_node_point(size=9, color='gold') +
  geom_node_text(aes(label=name))
```

```{r}
degrees = degree(G)
print("Degree Centrality:")
print(degrees)

```

## Closeness Centrality

Average distance to all other nodes

<div class="container">
<div class='fragment col'>
```{r}
G |>
  ggraph() +
  geom_edge_fan() +
  geom_node_point(size=9, color='gold') +
  geom_node_text(aes(label=name))
```
</div>

<div class='fragment col'>
```{r}
G |>
  mutate(closeness = centrality_closeness(normalized = T)) |>
  ggraph() +
  geom_edge_fan() +
  geom_node_point(aes(size=closeness), color='gold') +
  scale_size(range = c(2,10)) +
  geom_node_text(aes(label=name))
```
</div>


## Betweenness Centrality

Counts the number of shortest paths that go through each node. This is based on the value of being in a "structural hole"

Which of these has the highest betweenness?
<div class="container">
<div class='fragment col'>
```{r}
G |>
  ggraph() +
  geom_edge_fan() +
  geom_node_point(size=9, color='gold') +
  geom_node_text(aes(label=name))
```
</div>

<div class='fragment col'>
```{r}
G |>
  mutate(betweenness = centrality_betweenness(normalized = T)) |>
  ggraph() +
  geom_edge_fan() +
  geom_node_point(aes(size=betweenness), color='gold') +
  scale_size(range = c(2,10)) +
  geom_node_text(aes(label=name))
```
</div>


## Eigenvector Centrality

It's not what you know, it's who you know 

<div class="container">
<div class='fragment col'>
```{r}
G |>
  mutate(eigenvector_centrality = centrality_eigen()) |>
  ggraph() +
  geom_edge_fan() +
  geom_node_point(aes(size=eigenvector_centrality), color='gold') +
  scale_size(range = c(2,10)) +
  geom_node_text(aes(label=name))
```
</div>


# Network-level


## Density

Directed
$$
\frac{\sum(edges)}{N * (N-1)}
$$

Undirected: Multiply by 2



## Centralization

How unequal is the centrality between nodes?

<div class="container">
<div class='col'>
```{r, out.width='33%'}
set.seed(23)
x = random.graph.game(8, .3)

make_bw_plot <- function(G) {
  plot <- G |>
    as_tbl_graph() |>
  mutate(name = LETTERS[1:nrow(.N())]) |>
  mutate(betweenness = centrality_betweenness()) |>
  ggraph(layout='stress') +
  geom_edge_fan() +
  geom_node_point(aes(size=betweenness), color='gold') +
  scale_size(range = c(2,10)) +
  geom_node_text(aes(label=name))
  return(plot)
}

x |> make_bw_plot()

create_star(8, mode='undirected') |> make_bw_plot()
create_lattice(8, circular = T) |> make_bw_plot()
```
Betweenness Centralization scores:
```{r}
centralization.betweenness(x)$centralization
centralization.betweenness(create_star(8, mode='undirected'))$centralization
centralization.betweenness(create_lattice(8, circular = T))$centralization
```

## Hierarchy

Does information flow from the top down or in both directions?

```{r out.width='45%'}
library(sna)
library(intergraph)
library(networkdata)

ht_advice |>
  as_tbl_graph() |> 
  ggraph() +
  #geom_edge_fan(aes(alpha = stat(index))) + 
  #scale_edge_alpha_identity() +
  geom_edge_fan(arrow = arrow(length = unit(3, 'mm')), 
                 end_cap = circle(5, 'mm'),
                alpha = .7) + 
  geom_node_point(size = 9, color = 'gold')
  
print('Hierarchy of advice network:')
print(hierarchy(asNetwork(ht_advice)))
```
