---
pagetitle: COM 411
format: 
  revealjs:
    css: /resources/custom_css.css
    theme: white
    progress: true
    transition: convex
    reveal_plugins: ['notes']
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = './')
source("resources/preamble.R")
library(igraph)
library(tidygraph)
library(ggraph)
set_graph_style()
```


# Ego networks and network perceptions

## Sources of ego networks

> - Surveys of individuals
> - "Extraction" from complete network
> - Usually one step from ego


## Example - Full network

```{r}
#neigbhors(G, 1)
G = create_notable('zachary') |>
  activate(nodes) |>
  mutate(id = 1:nrow(.N()))

G = G|>
  mutate(dist_from_2 = as.vector(distances(G, 2))) |>
  mutate(is_neighbor = dist_from_2 <= 1)

plot = G |>
  ggraph(layout = 'lgl') + 
  geom_edge_fan(width=.5) + 
  theme(legend.position = 'right')

plot + geom_node_point(aes(), size = 3)

```

## Example - Ego network highlighted

```{r}
plot + geom_node_point(aes(color = as.factor(dist_from_2)), size = 3) + 
  theme(legend.position = 'right') + 
  labs(color = "Distance from ego")
```

## Example - Ego network

```{r}
G |>
  filter(is_neighbor == TRUE) |>
  ggraph(layout = 'lgl') +
  geom_edge_fan(width=.5) + 
  geom_node_point(aes(color = as.factor(dist_from_2)), size = 3) + 
  theme(legend.position = 'right') + 
  labs(color = "Distance from ego")
```



## What can we measure about ego networks?

> - Degree centrality (size)
> - Density
> - Groups
> - Distance
> - Homogeneity
> - Homophily
> - Composition

## How can we use ego network data?

> - GSS data
>	- Changes over time
>	- Comparison by location
> - Predicting individual outcomes
>	- Health
>	- Employment
>	- Happiness

# Network Perception

## Ego networks are the only part of the networks that we see

> - But are they real?
> - Perceived networks differ from observed networks (Bernard, Kilworth, and Sailer)
>	- "What people say about their communications bears no resemblance to their behavior" (BKS, 1982)
> - "Friendship" is often not mutual

## So what?

> - Perceptions are what influence actions
> - Potential network vs. "activated" network (Smith et al., 2012)
>	- Part of wealth and opportunity disparity may be due to differences in network activation

## How can we overcome perception biases?

> - Get information from others
>	- Build picture of whole network (Krackhardt et al)
> - Think in terms of potential networks and activation and mobilization
> - Recognize the value of "weak ties"



## Today's Dad Joke

I accidentally swallowed a bunch of scrabble tiles.

My next trip to the bathroom could spell disaster.

