---
title: Social Influence
format: 
  revealjs:
    css: /resources/custom_css.css
    theme: white
    progress: true
    transition: convex
    reveal_plugins: ['notes']
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = './')
source("resources/preamble.R")
set.seed(0121)
library(igraph)
library(tidygraph)
library(ggraph)
```
## Diffusion and Contagion

> - What we do is influenced by what others do
> - Network effects
>	- Standards, languages, technologies
> - Social constructs
>	- Concepts or things have meaning and importance through social 
> - Beliefs

## Simple contagion

> - Small world networks make simple contagion (like COVID-19) is really hard to stop!
> - Most interventions are network interventions
>	- Stopping travel attempts to block "long" ties
>	- Social distancing and shelter-in-place attempt to reduce degree
> - The most important people to influence are those with high betweenness centrality

## Example

<div class="r-stack">

<div class='fragment'>
```{r, message=F}
create_notable('zachary') |>
  mutate(index = 1:nrow(.N())) |>
  mutate(infected = index %in% c(2,3)) |>
  ggraph() +
  geom_edge_fan(width = .1) + 
  geom_node_point(aes(color = infected), size = 3) +
  theme_graph()
```
</div>

<div class='fragment'>
```{r, message=F}
create_notable('zachary') |>
  mutate(index = 1:nrow(.N())) |>
  mutate(infected = index %in% c(2,3)) |>
  mutate(neigh_2 = index %in% neighbors(.G(), 2),
         neigh_3 = index %in% neighbors(.G(), 3)) |>
  mutate(infected = infected | neigh_2 | neigh_3) |>
  ggraph() +
  geom_edge_fan(width = .1) + 
  geom_node_point(aes(color = infected), size = 3) +
  theme_graph()
```
</div>

</div>

## Complex contagion

> - Most communication-related diffusions are complex
>	- You don't change beliefs/behavior just from influence from one other person
> - Complex contagions require multiple ties between groups
>	- This is one reason groups can persist in different beliefs

## Example

<div class="r-stack">

<div class='fragment'>
```{r, message=F}
create_notable('zachary') |>
  mutate(index = 1:nrow(.N())) |>
  mutate(infected = index %in% c(2,3)) |>
  ggraph() +
  geom_edge_fan(width = .1) + 
  geom_node_point(aes(color = infected), size = 3) +
  theme_graph()
```
</div>

<div class='fragment'>
```{r, message=F}
create_notable('zachary') |>
  mutate(index = 1:nrow(.N())) |>
  mutate(infected = index %in% c(2,3)) |>
  mutate(neigh_2 = index %in% neighbors(.G(), 2),
         neigh_3 = index %in% neighbors(.G(), 3)) |>
  mutate(infected = infected | (neigh_2 & neigh_3)) |>
  ggraph() +
  geom_edge_fan(width = .1) + 
  geom_node_point(aes(color = infected), size = 3) +
  theme_graph()
```
</div>

</div>


## Special People

> - Why is the Mona Lisa the most famous painting in the world?
>	- Because there had to be a most famous painting -- success begets success (rich-get-richer)


## How does this apply to networks?

> - Can we predict who will trigger large-scale information cascades?
>	- Not very well
>	- If we look in hindsight, a viral tweet seems genius, appropriate, and well-timed
>	- There were likely hundreds or thousands of similar tweets
> - Luck + (ever-changing) structure of networks matter more than attributes of information or information spreaders
> - Network position is also closely related to luck!
