---
title: Scale-free networks and the friendship paradox
format: 
  revealjs:
    css: /resources/custom_css.css
    theme: white
    progress: true
    transition: convex
    reveal_plugins: ['notes']
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = './')
source("resources/preamble.R")
set.seed(0121)
library(igraph)
library(tidygraph)
library(ggraph)
library(tidyverse)
```

## Scale-free networks

> - What do degree distributions look like?
>	- In many cases, they are really skewed!
>	- A few people have many edges, while many have few
> - Why?
>	- Rich-get-richer processes (preferential attachment)

## Why are they called scale-free?

> - No matter where you zoom in on the distribution, it has the same shape

<div style="position:relative; width:640px; height:480px; margin:0 auto;">
<div class='fragment' style="position:absolute;top:0;left:0;">
```{r}
set_graph_style()

G = play_barabasi_albert(10000, growth = 1, power = 1)

df = G |> 
  as_tbl_graph() |>
  activate(nodes) |>
  mutate(degree = centrality_degree(mode='all')) |>
  as_tibble() |>
  group_by(degree) |>
  tally(name = 'count')


df |> 
  filter(degree > 1) |>
  ggplot(aes(x = degree, y = count)) + 
  geom_line() + 
  theme_minimal() +
  labs(title="All nodes")

```

</div>

<div class='fragment' style="position:absolute;top:0;left:0;">
```{r}
df |> 
  filter(degree > 2) |>
  ggplot(aes(x = degree, y = count)) + 
  geom_line() + 
  theme_minimal() + 
  labs(title="Nodes with degree of at least 3")
```
</div>

<div class='fragment' style="position:absolute;top:0;left:0;">
```{r}
df |> 
  filter(degree > 3) |>
  ggplot(aes(x = degree, y = count)) + 
  geom_line() + 
  theme_minimal() + 
  labs(title="Nodes with degree of at least 4")
```

</div>
</div>

## Implications

> - Robust to random failures
> - Outcomes are "unfair"


## Friendship paradox

> - Scott Feld (Purdue Sociologist!)
> - On average, your friends have more friends than you do

## What?!

> - The key idea is that people who are connected to lots of people are more likely to show up in your friend network.
> - You are likely to be friends with people who have lots of friends, because they have lots of friends!
> - This is much more pronounced as the skew of a network changes
> - E.g., the people you follow on Twitter/Instagram are likely much more popular than you

## Example

```{r, out.width="45%"}

set_graph_style()

G = barabasi.game(10, m = 2)

G = G |> as_tbl_graph() |>
  mutate(degree = centrality_degree(mode='all'),
         neighbor_degree = local_ave_degree())
p = G |> 
  ggraph(layout = 'kk') +
  geom_edge_fan() +
  geom_node_point(color='lightblue', size = 16) +
  geom_node_text(aes(label=degree), size = 12, fontface='bold')

p
```
<span class='fragment'>
```{r, out.width="45%"}

p + 
  geom_node_point(aes(color = neighbor_degree > degree), size = 16) +
  geom_node_text(aes(label=round(neighbor_degree, 2)), nudge_y = -.1, size=9) +
  geom_node_text(aes(label=degree), nudge_y = .1, size=9) +
  scale_color_viridis_d("Neighbors have higher degree", begin = .6, end = .85, )

```
</span>

## Implications

> - In many cases, we can't see who is most popular
> - Simply choosing a node at random, and then choosing one of their friends will identify a group that is more popular and closer to center
> - Detect outbreaks
> - Focus interventions
