---
title: Small worlds
format: 
  revealjs:
    css: /resources/custom_css.css
    theme: white
    progress: true
    transition: convex
    reveal_plugins: ['notes']
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = './')
source("resources/preamble.R")
set.seed(0121)
library(igraph)
library(tidygraph)
library(ggraph)
```

# What do human networks look like?

## Hard to measure for a long time

> - Erdos and Reyni came up with random graphs
> - Each edge has equal probability of existing
> - This obviously isn't true
> - Most of us are much more likely to be connected to those near us (along multiple dimensions)

## What do "groupy" networks look like?

> - High clustering (triadic closure)
> - Long path lengths
> - High "geodesic distance" (longest path between any two nodes)

## Milgram Experiment and 6 degrees

> - Try to get a letter to Mr. X
> - Found that (of those that made it) it took ~6 steps
  > - Similar results found in 2003 using email (Dodds, Muhamad, Watts)


## The puzzle

> - We know that we are in "groupy" networks yet all ~8 billion of us seem to be connected within a relatively short number of steps. How?

## Watts and Strogatz big insight

"Groupy" graph:

```{r, fig.height = 4.5, fig.width =4.5}
G = watts.strogatz.game(1, 40, 2, 0)

G |> ggraph('circle') + geom_edge_arc() + geom_node_point() + theme_graph()
```

Average distance: `r mean(distances(G), na.rm=T)`;
Max distance: `r max(distances(G), na.rm=T)`

## Just a few random "rewirings" make a clustered network with low distances

10% of edges rewired:

```{r, fig.height = 4.5, fig.width =4.5}
G = watts.strogatz.game(1, 40, 2, .1)

G |> ggraph('circle') + geom_edge_fan() + geom_node_point() + theme_graph()
```

Average distance: `r mean(distances(G), na.rm=T)`;
Max distance: `r max(distances(G), na.rm=T)`

## Where do these "rewirings" appear?

> - Hubs
> - People who move far from home
> - People whose interests diverge from those around them


## Lots of other networks have these same characteristics

> - Oracle of Bacon
> - Six Degrees of Wikipedia
