---
pagetitle: Tie formation
format: 
  revealjs:
    css: /resources/custom_css.css
    theme: white
    progress: true
    transition: convex
    reveal_plugins: ['notes']
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=F)
knitr::opts_knit$set(root.dir = './')
source("libs/preamble.R")
library(igraph)
library(tidygraph)
library(ggraph)
```

# Creation of edges

## Why do people form relationships / edges?

> - Psychological explanations
> - Strategic explanations
> - Exposure explanations

# Psychological explanations

## We are comfortable with people like us

> - Homophily
>   - People who are like each other are more likely to become friends

<div class='fragment'>
```{r, fig.height=3}
theme_set(theme_graph())
df = data.frame(from = c("A","B","C"), to = c("B","C","A"), type=c('enemy','enemy','friend'))
colors = data.frame(name=c("A","B","C"), color = c("gold", "black", "black"))

G = graph_from_data_frame(d = df, vertices = colors) |> as_tbl_graph()

plot = G |>
  activate(nodes) |>
  mutate(size = centrality_betweenness()) |>
  ggraph(layout = 'linear', circular= T) +
  #geom_node_point(aes(color = color), size = 8) + 
  geom_node_text(aes(label = name, color = color), size = 10) +
  scale_color_manual(values = c("gold"='gold',"black"='black'), guide = F) +
  #scale_edge_color_manual(values = c('enemy' = 'white', 'friend'='white'), guide = F)
  coord_cartesian(clip = 'off')
plot
```
</div>


## We want our relationships to be in balance

> - Balance theory
>   - The enemy of my enemy is my friend
>   - The friend of my enemy is my enemy

<div class='fragment'>
```{r, fig.height=3}
plot + scale_edge_color_manual(values = c('enemy'='black', 'friend'='red'), guide = F) +
geom_edge_fan(aes(label=type, color=type))
```
</div>
<aside class='notes'>
Balance theory is a bit more complicated. It says that we want our relationships to be in balance
If A is friends with  B and B is friends with C, but A and C are enemies, then balance theory says that this is unstable. Ties will dissolve or change valence.

This does depend on the type of relationship, though. What if black represents "hates"? If A hates B and B hates C, then maybe A will like C. Or if A is in love with B and B is in love with C, then A may hate C. These triangles are the source of lots of literature!
</aside>

# Strategic explanations

## We want to be in relationships that help us

> - Social exchange theory
>   - Do the benefits of relationship outweigh the costs?
>   - What resources does each node have and need?
> - Social capital
>   - People seek to make relationships with powerful others
>   - People seek to obtain powerful positions in the network

# Exposure explanations

## We become friends with our friends' friends

> - Triadic closure / transitivity

<div class='fragment'>
```{r, fig.height=3}

plot + geom_edge_fan(aes(color = type,
                         start_cap = label_rect(node1.name),
                         end_cap = label_rect(node1.name)),
                     width = .3, 
                     arrow = arrow(length = unit(3,'mm'))
				) + scale_edge_color_manual(values = c('enemy'='black', 'friend'='white'), guide = F)
           
```
</div>

<aside class="notes"> 
It is worth nothing that homophily is another explanation for the ubiquity of triadic closure. Those who are like us are likely to be our friends; those who are like our friends are likely to be like us.

</aside>


## Focus of Activity

> - Partially a pushback against other explanations
>	- People choose or are forced into foci (clubs, neighborhoods, classes, etc.)
>	- Ties are often a function of activities, not dyadic choices

