---
title: Networks in small groups
format: 
  revealjs:
    css: /resources/custom_css.css
    theme: white
    progress: true
    transition: convex
    reveal_plugins: ['notes']
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = './')
#source("resources/preamble.R")
library(igraph)
library(tidygraph)
library(ggraph)
library(patchwork)
```

## Much of our lives are lived in the context of small groups

> - Families
> - Class groups
> - Roommates
> - Friendship groups
> - Work groups
> - Volunteer organizations

## Lots of network theories relate to the structure of small groups

> - Triadic closure
> - Reciprocity
> - Homophily

## Research has also focused on what makes groups effective at meeting their goals

> - Which network do you think would be most effective?

<div class='fragment'>

```{r fig.width=8, fig.height=4}

star <- make_star(6) |>
  as_tbl_graph() |>
  ggraph(layout='fr') +
  geom_edge_fan() + 
  geom_node_point() + theme_void()

ring <- make_ring(6) |>
  as_tbl_graph() |>
  ggraph(layout='fr') +
  geom_edge_fan() + 
  geom_node_point() + theme_void()

complete <- make_full_graph(6) |>
  as_tbl_graph() |>
  ggraph(layout='fr') +
  geom_edge_fan() + 
  geom_node_point() + theme_void()

star + plot_spacer() + ring + plot_spacer() + complete

```

</div>

## It depends on the task!

> - Centralization is beneficial for simple tasks
> - Decentralization is beneficial for complex tasks or when information is scattered
> - Rings are pretty bad no matter what :)

## Communication isn't usually determined from on high

> - Communication networks both represent and enable patterns of communication

## Why do edges exist in small groups?

> - Social capital
> - Social exchange
> - Collective action
> - Transactive memory
>	- Know _who knows what_
>	- Knowledge is passed through communication

## In general, "integrative" groups perform better at complex tasks

> - Low hierarchy
> - Small periphery

> - In these groups, people are:
>	- Coordinated
>	- Motivated

## Groups are embedded

> - Work group vs. school assignment, etc.
> - People are also embedded in different groups
> - Tricky but important to consider the influence of environment

