---
title: COM 411
template: "../resources/template.html"
output: 
  revealjs::revealjs_presentation:
    theme: white
    progress: true
    transition: convex
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_knit$set(root.dir = './')
source("resources/preamble.R")

f <- function (x) {formatC(x, format="d", big.mark=',')}
bold <- function(x) {paste('{\\textbf{',x,'}}', sep ='')}
gray <- function(x) {paste('{\\textcolor{gray}{',x,'}}', sep ='')}
wrapify <- function (x) {paste("{", x, "}", sep="")}
p <- function (x) {formatC(x, format='f', digits=1, big.mark=',')}
library(igraph)
library(tidyverse)
```

<style type="text/css">

/* Changes to RevealJS default formatting.

I prefer this style but you can change these to your preferences */

/* This bit lets you do columns */
.container{
  display: flex;
}

.col{
  flex: 1;
}

.reveal section img{
border: none;
background: none;
box-shadow: none;
}

.highlight {
color: LightCoral;
}

.reveal h1,
.reveal h2,
.reveal h3,
.reveal h4,
.reveal h5,
.reveal h6 {
text-transform: none;
}

.reveal h1 {
margin-top:200px;
}

div.footnote {
font-size: 40%;
text-align: right;
}

.bg-image{
width: 180px;
}

.reveal .slides section .fragment.fade-in-then-out,
.reveal .slides section .fragment.current-visible {
  opacity: 0;
  visibility: hidden; }
  .reveal .slides section .fragment.fade-in-then-out.current-fragment,
  .reveal .slides section .fragment.current-visible.current-fragment {
    opacity: 1;
    visibility: inherit; }

.reveal .slides section .fragment.fade-in-then-semi-out {
  opacity: 0;
  visibility: hidden; }
  .reveal .slides section .fragment.fade-in-then-semi-out.visible {
    opacity: 0.5;
    visibility: inherit; }
  .reveal .slides section .fragment.fade-in-then-semi-out.current-fragment {
    opacity: 1;
visibility: inherit; }

</style>

## Today's Dad Joke

Just so everyone's clear,
<span class='fragment'>
I'm going to put my glasses on.
</span>

## Housekeeping

> - C-Span Archives Undergrad Research Competition
> - bit.ly/ccseurc20

# Troubled Lands Recap

## Main questions

> - How did you play differently in different versions?
> - What in real life is similar to each version of the game?
> - How did you decide whether or not to help others?
> - What does this have to do with social networks and social capital?!

## Debrief

> - What did you think about while playing?
> - How did you feel as you played? Did those feeling change across different versions of the game?
> - When do you help others in real life?

## Debrief

> - Were resources and abilities distributed fairly?
> - How should resources be distributed in the real world?
> - What role did communication play? How did your group make decisions?
> - Did you work to make things equal? Why or why not?
> - What kinds of inequality appeared in the game?

## Debrief

> - Did anyone sanction someone else? Why? 
> - How did others respond to being sanctioned?
> - What is similar to a "sanction" in the real world?

# Network Visualizations

## Aspects of visualizations

> - Nodes
> - Edges
> - Location

## Nodes

> - Information can be conveyed by:
>   - Shape, size, color
> - Shape
>   - Typically categorical (e.g., gender, age range)
> - Size
>   - Often a network measure, but can be something about node
> - Color
>    - Often community detection

## Examples

```{r}
#G = graph_from_data_frame(read.csv('../one_k_redittors.csv'))
G = random.graph.game(15, .4)
plot(G)
```

## With node shapes

```{r}
V(G)$shape = sample(c('square', 'circle','sphere'), size = gorder(G), replace = T)
plot(G)
ages = c('18-25', '26-34', '34-50')
```

## Nodes sized by betweenness centrality

```{r}
V(G)$shape = 'circle'
plot(G, vertex.size = betweenness(G) * 2.5)
```

## Nodes colored by community

```{r}

community = membership(fastgreedy.community(G))
colors = c('red', 'orange', 'darkgreen')
V(G)$color = colors[community]
plot(G)
```

## Edges

> - Size
>   - Typically represent weight of relationship
> - Color
>   - Typically represents different types of relationships

## Examples

## Edge width as weight
```{r}

make_edges = function(from, nodes = 15){
  x = 1:nodes
  x = x[x != from]
  to = sample(x, 1)
}

from = sample(1:15, size = 35, replace = T)
to = map_int(from, make_edges)
G = graph_from_edgelist(cbind(from,to))
E(G)$weight = sample(1:4, size = ecount(G), replace = T)
E(G)$type = sample(c(T,F), size=ecount(G), replace = T)
E(G)$arrow.size=.5

plot(G, edge.width = E(G)$weight)
```

## Edge color as type


```{r}
colors = c('gray', 'orange')
E(G)$color = colors[E(G)$type + 1]

plot(G)
```

## Position

> - Can represent literal distance
>   - Cities
>   - Seating chart
> - Or social distance
>   - Formal hierarchy
>   - Degree centrality

## Node placement can make a big difference in how a network is perceived

```{r, out.width='32%'}
G = random.graph.game(15, .4)
plot(G, layout = layout_in_circle(G))
plot(G, layout = layout_randomly)
plot(G, layout = layout_as_star)
plot(G)
plot(G, layout = layout_on_grid)
plot(G, layout = layout_with_lgl)
```

## Today's activity

> - Start working on DataCamp Network Analysis, chapters 2 and 3


# Network Analysis in R


## Today's Dad Joke

Within minutes, the detective knew exactly what the murder weapon was.

<span class='fragment'>
It was a brief case.
</span>

## Housekeeping

> - Self-assessment overview
>	- R
>	- Coming with questions / Office hours
>	- Readings
>	  - Too long
>	  - Put on Brightspace
>	- Posting slides
> - In class
>	- Break
>	- Dealing with confusion
>	  - More group discussion
>	  - Questions in class
>	  - Writing on projector vs. on board?


## DataCamp Review

> - neighbors()
> - ego()
> - make_ego_network()
> - Come back to:
>	- lapply
>	- randomization
>	- transitivity
>	- cliques


## What makes for a good network visualization?

Kozo Sugiyama's rules:

1. Lines should be straight.
2. Lines should be far apart from one another.
3. Lines should not cross or touch.
4. Lines should be easy to follow from one node to another.
5. Nodes that connect should be close.
6. Nodes that are most central should be close to the center of the graph.
7. Nodes that are similar in some way should be placed near one another.

## Activity

Find and Assess a network visualization

> - Use Google Images (or similar) to find an image of a network graphs.
> - What do you think the image does well and poorly? Does it follow Sugiyama's rules?
> - Put the image (and your assessment) on the padlet at https://padlet.com/jdfoote1/networks (link is also on wiki)

## Discussion

> - Did you agree with Sugiyama's rules?
> - What types of color/shape/size approaches were effective?

## Visualization Competition

> - Dutch school dataset (on wiki)
> - In groups of 2-3
>	- Come up with question
>	- Come up with visualization approach
> - We will work on this more next Thursday

## Homework
> - Scott Feld guest lecture
>	- Summary and questions assignment
