---
title: COM 411
template: "../resources/template.html"
output: 
  revealjs::revealjs_presentation:
    theme: white
    progress: true
    transition: convex
    self_contained: false
    reveal_plugins: ["notes"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_knit$set(root.dir = './')
source("resources/preamble.R")

f <- function (x) {formatC(x, format="d", big.mark=',')}
bold <- function(x) {paste('{\\textbf{',x,'}}', sep ='')}
gray <- function(x) {paste('{\\textcolor{gray}{',x,'}}', sep ='')}
wrapify <- function (x) {paste("{", x, "}", sep="")}
p <- function (x) {formatC(x, format='f', digits=1, big.mark=',')}
library(igraph)
```

<style type="text/css">

/* Changes to RevealJS default formatting.

I prefer this style but you can change these to your preferences */

/* This bit lets you do columns */
.container{
  display: flex;
}

.col{
  flex: 1;
}

.reveal section img{
border: none;
background: none;
box-shadow: none;
}

.highlight {
color: LightCoral;
}

.reveal h1,
.reveal h2,
.reveal h3,
.reveal h4,
.reveal h5,
.reveal h6 {
text-transform: none;
}

.reveal h1 {
margin-top:200px;
}

div.footnote {
font-size: 40%;
text-align: right;
}

.bg-image{
width: 180px;
}

.reveal .slides section .fragment.fade-in-then-out,
.reveal .slides section .fragment.current-visible {
  opacity: 0;
  visibility: hidden; }
  .reveal .slides section .fragment.fade-in-then-out.current-fragment,
  .reveal .slides section .fragment.current-visible.current-fragment {
    opacity: 1;
    visibility: inherit; }

.reveal .slides section .fragment.fade-in-then-semi-out {
  opacity: 0;
  visibility: hidden; }
  .reveal .slides section .fragment.fade-in-then-semi-out.visible {
    opacity: 0.5;
    visibility: inherit; }
  .reveal .slides section .fragment.fade-in-then-semi-out.current-fragment {
    opacity: 1;
visibility: inherit; }

</style>

## Today's Dad Joke
Bro, can you pass me that leaflet?


Brochure.

## Housekeeping

> - Thursday Dr. Lee will be visiting
> - Next Thursday is an in-class exam

## Reflection summary

> - Good
>	- Activities
>	- Class discussion
> - Bad
>	- Six Degrees book
>	  - Sorry!
>	- Brightspace + wiki
>	  - Not sorry ;)
> - Confusing
>	- What kinds of goals to set?

## What to know from Six Degrees

> - How closely connected are people who are geographically/socially distant?
> - People are clustered: our friends are friends with each other, so it feels like it should be distant
>	- High clustering coefficient
> - Human networks have short diameter
>	- We are all connected within a surprisingly small distance


## Homework Review

>  - Questions about networks in R?

# Power in Social Networks

## What is power?
> - Long, varied history in the social sciences
>   - One version is the capacity to achieve one's will
>   - How do we do that? Through capital
> - Social networks are one way of measuring "social capital"
> - Who has power in a social network?

## Degree Centrality

>- Counts the number of edges each node has
>  - In, out, or all

```{r, out.width="50%"}
set.seed(23)
G = random.graph.game(8, .35)

plot(G, vertex.size = (degree(G) +1)* 4)

degrees = degree(G)
names(degrees) = 1:8
print("Degree Centrality:")
print(degrees)

```

## Closeness Centrality

Average distance to all other nodes

<div class="container">
<div class='fragment col'>
```{r}
plot(G)
```
</div>

<div class='fragment col'>
```{r}
plot(G, vertex.size = closeness(G, normalized = T) * 50)
```
</div>


## Betweenness Centrality

Counts the number of shortest paths that go through each node. This is based on the value of being in a "structural hole"

Which of these has the highest betweenness?
<div class="container">
<div class='fragment col'>
```{r}
plot(G)
```
</div>

<div class='fragment col'>
```{r}
plot(G, vertex.size = betweenness(G) * 5)
```
</div>


## Eigenvector Centrality

Who is connected to the most important people? If you have only one friend, but that friend is the President, then you are still powerful.

<div class="container">
<div class='fragment col'>
```{r}
plot(G, vertex.size = eigen_centrality(G)$vector * 40)
```
</div>

<div class='fragment col'>
```{r}
pa = barabasi.game(8, directed = F)
plot(pa,vertex.size = eigen_centrality(pa)$vector * 40)
```
</div>

## Network-level


## Density

Directed
$$
\frac{\sum(edges)}{N * (N-1)}
$$

Undirected: Multiply by 2



## Centralization

How unequal is the centrality between nodes?

<div class="container">
<div class='col'>
```{r, out.width='33%'}
x = random.graph.game(8, .3)
plot(x)

plot(make_star(8, mode='undirected'))

plot(make_lattice(8, circular = T))
```

## Using metadata to find Paul Revere

> - Where did the data come from?
>    - What kind of network was it?
> - Did this make you think differently about metadata? 
>	- About privacy?


## Homework

> - Reading
> - Keep working on R
