---
pagetitle: COM 411
output: 
  revealjs::revealjs_presentation:
    theme: white
    progress: true
    transition: convex
    self_contained: false
    css: ../../libs/custom_css.css
    reveal_plugins: ["notes"]
    lib_dir: ../../libs
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = './')
source("resources/preamble.R")
library(igraph)
library(tidygraph)
library(ggraph)
```


# Ego networks

## Sources of ego networks

> - Surveys of individuals
> - "Extraction" from complete network
> - Usually one step from ego


## Example - Full network

```{r}

G = create_notable('karate')

```



## What can we measure about ego networks?

> - Degree centrality (size)
> - Density
> - Groups
> - Distance
> - Homogeneity
> - Homophily
> - Composition

## How can we use ego network data?

> - GSS data
>	- Changes over time
>	- Comparison by location
> - Predicting individual outcomes
>	- Health
>	- Employment
>	- Happiness



## Today's Dad Joke
I accidentally swallowed a bunch of scrabble tiles.

My next trip to the bathroom could spell disaster.

