---
title: Diffusion and Contagion
template: "../resources/template.html"
output: 
  revealjs::revealjs_presentation:
    theme: white
    progress: true
    transition: convex
    self_contained: false
    css: ../../libs/custom_css.css
    reveal_plugins: ["notes"]
    lib_dir: ../../libs
    allow_uptree_lib_dir: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = './')
source("resources/preamble.R")
set.seed(0121)
library(igraph)
library(tidygraph)
library(ggraph)
```

## Dad Joke

Interviewer: "Where do you see yourself in 5 years?"

<span class='fragment'>
Me: "Listening. I would say listening is my biggest weakness."
</span>

## Housekeeping

> - Sorry about due date for Dutch School assignment!!
> - Self-assessment reflections due today
>	- Feedback by next week
> - Troubled Lands game on Thursday

## Upcoming Assignments

> - Exam
>	- Visualization + Essays
>	- Submit two questions for short answer section
> - Groups in networks assignment

## Example questions
> - How does one gain social capital from other forms of capital? Give three examples with explanations.
> - What does "centrality" measure? How do "betweenness centrality", "degree centrality", and "eigenvector centrality" differ, and when might one be a better measure than the other?
> - What is the “Friendship Paradox”?  How does it work? How does it relate to information flow and contagion?
> - What is a "multiplex" (or multidimensional) network? Give an example where thinking of a network as multiplex could help you to understand it better.

## Plan for today

> - Diffusion and contagion review
> - Visualization discussion + show and tell
> - Discussion questions

## Review Questions

> - What are some things that are diffused through a social network?
> - What is the difference between simple contagion and complex contagion?

## Simple Contagion

<div class="r-stack">

<div class='fragment'>
```{r, message=F}
create_notable('zachary') %>%
  mutate(index = 1:nrow(.N())) %>%
  mutate(infected = index %in% c(2,3)) %>%
  ggraph() +
  geom_edge_fan(width = .1) + 
  geom_node_point(aes(color = infected), size = 3) +
  theme_graph()
```
</div>

<div class='fragment'>
```{r, message=F}
create_notable('zachary') %>%
  mutate(index = 1:nrow(.N())) %>%
  mutate(infected = index %in% c(2,3)) %>%
  mutate(neigh_2 = index %in% neighbors(.G(), 2),
         neigh_3 = index %in% neighbors(.G(), 3)) %>%
  mutate(infected = infected | neigh_2 | neigh_3) %>%
  ggraph() +
  geom_edge_fan(width = .1) + 
  geom_node_point(aes(color = infected), size = 3) +
  theme_graph()
```
</div>

</div>


## Complex Contagion

<div class="r-stack">

<div class='fragment'>
```{r, message=F}
create_notable('zachary') %>%
  mutate(index = 1:nrow(.N())) %>%
  mutate(infected = index %in% c(2,3)) %>%
  ggraph() +
  geom_edge_fan(width = .1) + 
  geom_node_point(aes(color = infected), size = 3) +
  theme_graph()
```
</div>

<div class='fragment'>
```{r, message=F}
create_notable('zachary') %>%
  mutate(index = 1:nrow(.N())) %>%
  mutate(infected = index %in% c(2,3)) %>%
  mutate(neigh_2 = index %in% neighbors(.G(), 2),
         neigh_3 = index %in% neighbors(.G(), 3)) %>%
  mutate(infected = infected | (neigh_2 & neigh_3)) %>%
  ggraph() +
  geom_edge_fan(width = .1) + 
  geom_node_point(aes(color = infected), size = 3) +
  theme_graph()
```
</div>

</div>

## Review Questions

> - What is an example of simple contagion?
> - What is an example of something that spreads via complex contagion?
> - What was the argument of the "Special People" chapter?

## R Show and Tell

## Discussion Questions

