---
title: COM 411
template: "../resources/template.html"
output: 
  revealjs::revealjs_presentation:
    theme: white
    progress: true
    transition: convex
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_knit$set(root.dir = './')
source("resources/preamble.R")

f <- function (x) {formatC(x, format="d", big.mark=',')}
bold <- function(x) {paste('{\\textbf{',x,'}}', sep ='')}
gray <- function(x) {paste('{\\textcolor{gray}{',x,'}}', sep ='')}
wrapify <- function (x) {paste("{", x, "}", sep="")}
p <- function (x) {formatC(x, format='f', digits=1, big.mark=',')}
library(igraph)
library(tidyverse)
```

<style type="text/css">

/* Changes to RevealJS default formatting.

I prefer this style but you can change these to your preferences */

/* This bit lets you do columns */
.container{
  display: flex;
}

.col{
  flex: 1;
}

.reveal section img{
border: none;
background: none;
box-shadow: none;
}

.highlight {
color: LightCoral;
}

.reveal h1,
.reveal h2,
.reveal h3,
.reveal h4,
.reveal h5,
.reveal h6 {
text-transform: none;
}

.reveal h1 {
margin-top:200px;
}

div.footnote {
font-size: 40%;
text-align: right;
}

.bg-image{
width: 180px;
}

.reveal .slides section .fragment.fade-in-then-out,
.reveal .slides section .fragment.current-visible {
  opacity: 0;
  visibility: hidden; }
  .reveal .slides section .fragment.fade-in-then-out.current-fragment,
  .reveal .slides section .fragment.current-visible.current-fragment {
    opacity: 1;
    visibility: inherit; }

.reveal .slides section .fragment.fade-in-then-semi-out {
  opacity: 0;
  visibility: hidden; }
  .reveal .slides section .fragment.fade-in-then-semi-out.visible {
    opacity: 0.5;
    visibility: inherit; }
  .reveal .slides section .fragment.fade-in-then-semi-out.current-fragment {
    opacity: 1;
visibility: inherit; }

</style>


## Today's Dad Joke

I had a bunch of books fall on me.

I have only my shelf to blame.


## Housekeeping

> - Will respond soon to Self Assessment (and other assignments)
> - Keep working on Final Project
> - Reading summaries + FlipGrid

## Programming

> - Maybe turning a corner?
> - Reach out if you are still stuck!!


## The dark side of networks

> - Network measurement issues
> - Unintended consequences
> - Ethics

## Measurement issues

> - Network analysis is about identifying and considering interdependence
> - This is the biggest strength and the biggest weakness
> - Some edges and nodes are very important to a true understanding of the network

## Centralization of preferential attachment graph is high

```{r}
library(igraph)
G = as.undirected(barabasi.game(50, m = 2))

plot(G)
centralization.degree(G)$centralization
```
## What if we remove the most central node?

```{r}
G1 = delete_vertices(G, which(degree(G)==max(degree(G))))
plot(G1)
centralization.degree(G1)$centralization
```
## Vs a random node

```{r}
G2 = delete_vertices(G, sample(V(G), 1))
plot(G2)
centralization.degree(G2)$centralization
```

## Similar dynamics occur for many graph-level and individual-level measures

> - This is important because data is often missing but missing data is more important than when measurements are independent

## Networks can have unintended consequences

> - Globalized networks allow for
>    - Specialization
>    - Opportunity to enter global markets
>    - Pandemics
> - Computer networks allow for
>   - Unfettered communication
>   - viruses
>   - ransomware

## Network analysis can be used for ill

> - Belief diffusion can also be wrong beliefs
> - Propagandists can identify the vulnerable
> - Misinformation can spread as quickly as truth
> - Key hubs can be identified (power grid, etc.)
